\chapter{Larger Large Cardinals}

\section{The Measure Problem}

This section will motivate the definition of a Measurable Cardinal and state some equivalences that will be useful in further investigations later on the notes.

\subsection{Set Theoretic Version}

It's a classical problem to seek for a non-constant, translational invariant, $\sigma$-additive probabilistic measure $\mu:\mc{P}(\mbb{R})\to[0,1]$ in the reals $\mbb{R}$. And, as we know, Vitali showed such task is impossible.

The construction is rather simple for a modern set theorist, consider the real numbers modulo a rational, $\mbb{R}/\mbb{Q}$, and let $V$ be a transversal in $[0,1]$. We will show
$$[0,1]\subseteq\bigcup_{q\in[-1,1]\cap\mbb{Q}}V+q\subseteq[-1,2].$$
Since $V\subseteq[0,1]$, the second inclusion is clear. For the first let $r\in[0,1]$, then by definition of $V$ we can find $v\in[r]\cap V$, so $r-v\in[-1,1]\cap\mbb{Q}$, and then $r=v+(r-v)\in V+(r-v)$. So
$$0<\mu([0,1])\leq\sum_{n}\mu(V)\leq\mu([-1,2])<1$$
which is impossible for any value $\mu(V)$ could assume.

From here, there are two possible paths. Measure theory lead the one weakening the domain of $\mu$, i.e., considering the measure only for the sets in a $\sigma$-algebra of $\mbb{R}$. Set theory lead the other where the geometrical condition of translational invariance is weakened.

Banach showed that, without translational invariance, we could find trivial measures $\mu_r$ on any set $S$, with $r\in S$, as follows:
\begin{equation}\label{eq:banach_trivial_measure}
    \mu_r(A)=\begin{cases}
    1,\text{ if }r\in A;\\
    0,\text{ otherwise}.
\end{cases}
\end{equation}
So he imposed that, instead of translational invariance, we had a non-triviality condition, i.e., that $\mu(\{x\})=0$, $\forall x\in S$ which, with $\sigma$-additivity, guarantees us that all finite sets have measure $0$ too.

So, for the sake of our sanity, let's just pin down that, throughout this section, a function $\mu:\mc{P}(S)\to[0,1]$ is a measure on/over $S$ if $\mu(\emptyset)=0$, $\mu(S)=1$, it's non-trivial and it's $\sigma$-additive.

A measure on $S$ depends only on $\lvert S\rvert$, so for now on we will consider measures solely on cardinals. We can then state the set-theoretic version of the Measure Problem, as proposed by Banach:

\begin{tcolorbox}[colback=gray!5,colframe=gray!80!black]
\centering
\lasereye{\textbf{Set Theoretic Measure Problem (Banach)}}

Can we find a measure on $\mf{c}$. Or rather at least on some infinite $\kappa$?
\end{tcolorbox}

In order to keep investigating measures on infinite cardinals, consider the following generalization of the notion of $\sigma$-additiveness, which guarantees us that, as for the finite sets, any set of cardinality $\lambda<\kappa$ has also measure zero.

\begin{definition}{$\kappa$-additive measure}{kappa_additive_measure}
    We say a measure $\mu$ is \lasereye{$\kappa$-additive} if, for any $\lambda<\kappa$ and pairwise disjoint sets $\{X_\alpha\}_{\alpha<\lambda}$,
    $$\mu\rp{\bigcup_{\alpha<\lambda}X_\alpha}=\sum_{\alpha<\lambda}\mu(X_\alpha):=\sup\left\{\sum_{\alpha\in E}\mu(X_\alpha):E\in[\lambda]^{<\omega}\right\}.$$
\end{definition}

With such nomenclature, $\sigma$-additive is just $\aleph_1$-additive.

In fact, if $\mu$ is a $\kappa$-additive measure on a set $S$, then any subset of size $<\kappa$ has measure $0$, as desired. The next lemma explains why we care about $\kappa$-additiveness, it says that the solution for the Set Theoretic Measure Problem is related to the existence of a $\kappa$-additive measure.

\begin{lemma}{}{measure_implies_kappa_add}
    Let $\kappa$ be the least cardinal with a measure in it, then every measure over $\kappa$ is $\kappa$-additive.
\end{lemma}

\begin{proof}
    Assume by a contradiction that there are pairwise disjoint $\{X_\alpha\}_{\alpha<\lambda}$ with
    \begin{equation}\label{eq:auxiliary2}
        \mu\rp{\bigcup_{\alpha<\lambda}X_\alpha}\neq\sum_{\alpha<\lambda}\mu(X_\alpha).
    \end{equation}
    As I said, $\mu(X_\alpha)\neq0$ for at most countably many $\alpha$'s, so assume WLOG (by removing and reindexing them if necessary) that $\mu(X_\alpha)=0$ for any $\alpha<\lambda$. But, by eq. (\ref{eq:auxiliary2}), we have
    $$\mu\rp{\bigcup_{\alpha<\lambda}X_\alpha}=r>0.$$
    So the following $\widetilde{\mu}:\mc{P}(\lambda)\to[0,1]$ is a measure over $\lambda$:
    $$\widetilde{\mu}(Y):=\frac1r\cdot \mu\rp{\bigcup_{\alpha\in A}X_\alpha}$$
    contradicting the minimality of $\kappa$.
\end{proof}

So, now that we saw the importance of $\kappa$-additive measures, we can then define:

\begin{definition}{Real-Valued Measurable}{real_valued_measurable}
    A cardinal $\kappa>\omega$ is \lasereye{real-valued measurable} if there is a nontrivial $\kappa$-additive measure on $\kappa$.
\end{definition}

Since a solution for the Set Theoretic Measure Problem implies the existence of a real-valued measurable cardinal, we can now ask if there is any real-valued measurable cardinal.

As expected of a Large Cardinal Notes, the answer for this problem is related to the existence of some large cardinals, and therefore is independent of ZFC. To see this, note that, if $\kappa$ is real-valued measurable, by $\kappa$-additivity we have that $\kappa$ is regular. In fact, Banach established that, under GCH, every real-valued measurable cardinal is weakly inaccessible. But we can do even better! Ulam showed in his doctoral dissertation that we can remove GCH from Banach's result, as we will see later.

\subsection{Measures with Atoms}

Before proving the previous results, let's remark an important dichotomy between measures:

\begin{definition}{Atoms}{}
    Let $\kappa$ be real-valued measurable. A subset $A\subseteq\kappa$ with $\mu(A)>0$ is called an \lasereye{atom} if for every $B\subseteq A$, either $\mu(B)=0$ or $\mu(B)=\mu(A)$.

    We say a measure $\mu$ over $\kappa$ is \lasereye{atomless} if there is no atom $A\subseteq\kappa$.
\end{definition}

What is so important about being or not atomless? Note that, if $\mu$ has an atom $A\subseteq\kappa$, then
$$\widetilde{\mu}(X)=\frac{\mu(A\cap X)}{\mu(A)}$$
is a $2$-valued measure. We clearly have a correspondence, since every $A$ with positive measure is an atom in a $2$-valued measure. So, in some sense, atoms gives us some sort of generalization of Banach's trivial measure (\ref{eq:banach_trivial_measure}). We will see, however, that such measures are far from being trivial, and, maybe unexpectedly, will be way more fruitful to us than atomless measures!

As a matter of comparison, you could ask yourself if atomless measure can be as trivial as $2$-measures; for example, if it could be a $3$-measure. Fortunately, it's a theorem of Sierpiński that, if $\mu$ is atomless, given $X\subseteq S$ and $0\leq\alpha\leq\mu(X)$, there is some set $Y\subseteq X$ such that $\mu(Y)=\alpha$. In other words, $\mu$ attains all possible values it could possibly reach.

So, the reason why atoms are such a big deal is that we can make a correspondence between $\kappa$-complete free ultrafilters on $\kappa$ and $\kappa$-additive non-trivial measures with atoms on $\kappa$, as follows:
\begin{itemize}
    \item Given an ultrafilter $U$, let $\mu(X)=1$ iff $X\in U$;
    \item Given a measure $\mu$, let $X\in U$ iff $\mu(X)=1$.
\end{itemize}

Which gives rise to the following Large Cardinal Axiom.

\begin{definition}{Measurable Cardinal}{measurable_cardinal}
    A cardinal $\kappa$ is \lasereye{measurable} (denoted $M(\kappa)$) if there is a $2$-valued $\kappa$-additive measure on it.

    Equivalently, $\kappa$ is measurable if there is a $\kappa$-complete free ultrafilter on it.
\end{definition}

As noted earlier, $\kappa$-additivity implies regularity of $\kappa$. In fact, with atoms we can prove even more!

\begin{theorem}{}{measurable_inaccessible}
    Every measurable cardinal is inaccessible.
\end{theorem}

\begin{proof}
    Let $\kappa$ be a measurable cardinal and $U$ be its $\kappa$-complete free ultrafilter.
    
    Assume by a contradiction that there is some $\lambda<\kappa$, with $2^\lambda\geq\kappa$. Then we can find an injection $F:\kappa\to2^\lambda$ such that $F_\alpha:\lambda\to2$. So, given $\xi\in\lambda$, we have
    $$\kappa=\{\alpha\in\kappa:F_\alpha(\xi)=0\}\sqcup\{\alpha\in\kappa:F_\alpha(\xi)=0\}.$$
    But, since $U$ is an ultrafilter, either one of these sets are in $U$. Let $b_\xi$ be such that
    $$A_\xi=\{\alpha\in\kappa:F_\alpha(\xi)=b_\xi\}\in U.$$
    Then, by $\kappa$-completeness of $U$, we have that
    $$A=\bigcap_{\xi<\lambda}A_\xi\in U.$$
    However, if $\alpha\in A$, then $F_\alpha(\xi)=b_\xi$ for every $\xi\in\lambda$, i.e., $F_\alpha$ is a completely determined function, so $\lvert A\rvert\leq 1$, contradicting that $U$ is free and $A\in U$.
\end{proof}

This guarantees us that measurability is a Large Cardinal Axiom. Whether this is equivalent or not to inaccessibility was an open problem for quite some time. As expected (otherwise we weren't spent a whole section talking about measurables) they're not equivalent, which we will prove later after introducing some other large cardinals axioms.

\subsection{Real-Valued Measures}

Finally, as promised, let's return to the study of real-valued measures and see how the existence of atomless measures gives us some weird results about the continuum.

The previous subsection shows us that there is no hope to find\footnote{at least in ZFC} a $2$-valued measure on $\mf{c}$, but what about an atomless one? As we spoiled earlier, real-valued measurability implies weakly inaccessibility, which is now clear for the case with atoms. In order to prove this for real-valued measures, we can't use the set of subsets with measure $1$, since it won't be a filter anymore.

So, to do what we want, let's analyze it's dual notion! I.e., the ideal of null subsets of $\kappa$:
$$I_\mu:=\{X\subseteq\kappa:\mu(X)=0\}.$$
It's clear that, if $\mu$ is a $\lambda$-additive measure, then $I_\mu$ is a free $\lambda$-complete ideal. Morever, it's what we call a $\sigma$-saturated ideal, i.e., every family $\mc{F}$ of pairwise disjoint subsets of $\kappa$ that aren't in $I_\mu$ is at most countable. To see this, note that
$$\mc{F}=\bigcup_{n\geq1}\{X\in\mc{F}:\mu(X)\geq1/n\}$$
where each term in the RHS is at most finite.

This may seem like a not really powerful remark, but the fact that the ideal of null sets is $\sigma$-saturated is actually really strong! It's the reason why a lot of the following beautiful results work and also why we can relate the Set Theoretic Measure Problem with measures on $\mbb{R}$. In fact, note that $\sigma$-saturation follows solely by the fact that we decided to measure sets with reals\footnote{that's why it's called a real-valued measure, after all. despite $\kappa$-additivity, one cold argue that such generalization is not a completely set-theoretic one, since it relies on measuring as ``humans'' naturally would} in $[0,1]$.

About $\lambda$-additivity implying $\lambda$-completeness, we can also prove the converse result:

\begin{lemma}{}{completeness_additivity}
    Let $\mu$ be a measure on $\kappa$, if $I_\mu$ is $\lambda$-complete, then $\mu$ is $\lambda$-additive.
\end{lemma}

\begin{proof}
    Given $\gamma<\lambda$, let $\{X_\alpha\}_{\alpha<\gamma}$ be a family of pairwise disjoint subsets of $\kappa$, since our ideal $I_\mu$ is $\sigma$-saturated, we know at most countably many of those $X_\alpha$'s have positive measure, let's say $Y_n$, for $n\in\omega$. But then, by $\kappa$-completeness of $I_\mu$ and $\sigma$-additivity of $\mu$:
    $$\mu\rp{\bigcup_{\alpha<\gamma}X_\alpha}=\mu\rp{\bigcup_{n\in\omega}Y_n}=\sum_{n\in\omega}\mu(Y_n)=\sum_{\alpha<\gamma}\mu(X_\alpha)$$
    as desired. So $\mu$ is $\kappa$-additive.
\end{proof}

Let's now prove our first consequence about the existence of atomless measures.

\begin{theorem}{}{atomless}
    If there is an atomless measure $\mu$, then there is also one on some $\kappa\leq\mf{c}$.
\end{theorem}

\begin{sketch}
    Let's first sketch this proof, because it's really nice. It makes clear how different atomless measures are and how ``strong'' the $\sigma$-saturation property is.

    Let $\mu$ be such measure on $S$. The fact that it's atomless gives us the following amazing property that resembles disconnectedness:
    $$\text{If $X\subseteq S$, then we can write $X=Y\sqcup Z$, where $\mu(Y),\mu(Z)>0$}$$
    I.e., we can partition any set in subsets of positive measure. Otherwise $X$ would be an atom.

    Now, starting with $S$, we can partition it in subsets of positive measure, and then partition each of these subsets in other subsets of positive measure. If we keep going on such process, we end up with an infinite tree, where in the limit level we take all possible intersections of transversal with positive measure. 
    
    There is actually more than one way to construct such tree, we will explicit just one construction, thought. Let's now analyze how such tree looks like.

    Each level of our tree consists of pairwise disjoint sets of positive measure, then, by $\sigma$-saturation, each level has countably many sets. What about the size $\gamma$ of our tree? Let 
    $$b=\{X_\alpha:\alpha<\lambda\},\text{ with }\lambda<\gamma\text{ and }X_\alpha\text{ a set in the $\alpha$-th level}$$
    be branch of it. By definition, $\mu(X_{\alpha+1})<\mu(X_\alpha)$, and so $Y_\alpha=X_{\alpha+1}\setminus X_\alpha$ form a family of $\lambda$-many pairwise disjoint sets of positive measure. Again, by $\sigma$-saturation, we have that the size of each branch is at most countable, and so the size $\gamma$ of our tree is at most $\aleph_1$.

    Now, look at the limit point of each branch, they all have measure 0, otherwise it wouldn't be the limit point, since we could keep going on our process. They also form a partition of $S$ and, since it's a tree whose branches has countable height, and each level has countably many elements, there are at most $\mf{c}$ of these limit points.

    So we've constructed $\leq\!\mf{c}$-many null sets, whose union is $S$, i.e., our measure isn't $\mf{c}^+$-additive! Then it just remains to induce a measure on a smaller cardinal.
\end{sketch}

\begin{proof}
    For this first construction we will make some kind of binary tree $T$. Start with $S$ and, if the $\alpha$-th level is already defined, given some set $X$ in it, its immediate successors are the sets $Y$ and $Z$ of positive measure that partition $X$. And when $\alpha$ is a limit ordinal, we get all possible intersections $\bigcap_{\beta<\alpha}X_\beta$ with positive measure, where $X_\beta$ is in the $\beta$-th level.

    If it wasn't clear from the sketch, each branch has countable length in $T$, whose height is at most $\aleph_1$, so we can associate it branch with a countable subset of $T$, i.e., the amount $\kappa$ of branches is st
    $$\kappa\leq\left\lvert[T]^{\aleph_0}\right\rvert\leq\aleph_1^{\aleph_0}\leq\rp{2^{\aleph_0}}^{\aleph_0}=2^{\aleph_0\cdot\aleph_0}=\mf{c}$$
    Let then $\{b_\alpha:\alpha<\kappa\}$ be an enumeration of all branches $b_\alpha=\{X_\beta:\beta<\gamma\}$ whose limit point
    $$Z_\alpha:=\bigcap_{\beta<\gamma}X_\beta$$
    is nonempty. Then $\{Z_\alpha:\alpha<\kappa\}$ partition $S$ in $\kappa$ sets of measure $0$.
    
    We then induce a measure $\nu$ on $\kappa$ as follows: Let $f:S\to\kappa$ be defined as
    $$f(x)=\alpha\iff x\in Z_\alpha$$
    and let
    $$\nu(X):=\mu(f^{-1}(X)).$$
    It's not hard to verify that $\nu$ is in fact a measure on $\kappa$. To see it can't have an atom, assume it has, then $\kappa$ is measurable and, by \cref{th:measurable_inaccessible}, it would be inaccessible. However, you can't have an inaccessible bellow the continuum, otherwise, since $\aleph_0<\kappa\leq2^{\aleph_0}$, then $2^{\aleph_0}<\kappa\leq2^{\aleph_0}$.
\end{proof}

Clearly we can extend $\nu$ to a measure on $\mf{c}$, just let $\widetilde{\nu}(X)=\nu(X\cap\kappa)$, thus we conclude there is an atomless measure on $\mbb{R}$. It turns out we can, with a slightly modification of the previous argument, obtain a measure on $\mbb{R}$ extending the Lebesgue measure!

Such construction is done by partition the sets in a finer way, where we let $X_\emptyset=S$, and
$$X_{s\frown0}\sqcup X_{s\frown1}=X_s,\text{ with }\mu(X_{s\frown0})=\mu(X_{s\frown1})=\frac12\mu(X_s)$$
for every binary sequence $s$. We then let
$$\widetilde{\lambda}(X)=\mu\rp{\bigcup\{X_s:s\in X\}},\text{ where }X_s=\bigcap_{n\in\omega}X_{s\restriction_n}$$
with $\mbb{R}$ being identified as the Cantor Space $2^\omega$. The work now is to verify that $\widetilde{\lambda}$ is indeed a measure, and that it agrees with the Lebesgue Measure $\lambda$ on all intervals $[k/2^n,(k+1)/2^n]$, and hence on all Borel sets.

It now remains to show Ulam's main result, that every real-valued measurable cardinal is weakly inaccessible. In order to do this, we will need the following object:

\begin{definition}{Ulam Matrix}{ulam_matrix}
    For any cardinal $\lambda$, an \lasereye{Ulam $(\lambda^+,\lambda)$-matrix} is a collection $\{A_{\alpha,\eta}:\alpha<\lambda^+,\eta<\lambda\}$
    of subsets of $\lambda^+$ such that:
    \begin{itemize}
        \item $A_{\alpha,\eta}\cap A_{\beta,\eta}=\emptyset$;
        \item $\lvert\lambda^+\setminus\bigcup_{\eta<\lambda}A_{\alpha,\eta}\rvert\leq\lambda$, for each $\alpha<\lambda^+$.
    \end{itemize}
\end{definition}

In order to use it, let's guarantee that at least one Ulam $(\lambda^+,\lambda)$-Matrix.

\begin{lemma}{}{ulam_matrix}
    For any $\lambda$, there is an Ulam $(\lambda^+,\lambda)$-matrix.
\end{lemma}

\begin{proof}
    Given $\xi<\lambda^+$, let $f_\xi$ be function on $\lambda$ with $\xi\subseteq\text{Im}(f_\xi)$. Let then define, for $\alpha<\lambda^+$, $\eta<\lambda$,
    $$\xi\in A_{\alpha,\eta}\iff f_\xi(\eta)=\alpha$$
    Then, if $\alpha\neq\beta$, clearly $f_\xi(\eta)$ can't be equal both, so $A_{\alpha,\eta}\cap A_{\beta,\eta}=\emptyset$. Now, given $\alpha<\lambda^+$, let $\xi>\alpha$, so, since $\xi\subseteq\text{Im}(f_\xi)$, there is some $\eta<\lambda$ such that $f_\xi(\eta)=\alpha$, i.e., $\xi\in A_{\alpha,\eta}$, therefore
    $$\xi\in\bigcup_{\eta<\lambda}A_{\alpha,\eta},\text{ for all }\xi>\alpha$$
    whose complement is contained in $\alpha\leq\lambda$, as desired.
\end{proof}

Let's see how Ulam Matrices breaks $\sigma$-saturation.

\begin{theorem}{}{real_valued_weakly_inacc}
    Every real-valued measurable cardinal is weakly inaccessible.
\end{theorem}

\begin{sketch}
    Let's do the proof for $\lambda=\omega_1$, showing there can't be a measure on it.

    Intuitively, an Ulam $(\omega_1,\omega)$-matrix is a $\omega_1\times\omega$ matrix whose entries are subsets of $\omega_1$, positioned in such a way that:
    \begin{itemize}
        \item any two elements in a column are disjoint;
        \item the union of all elements in a row contains all, but countably many elements of $\omega_1$.
    \end{itemize}
    In fact, this is exactly the combinatorial object we want to break $\sigma$-saturation! Since the union of all elements in a row is $\omega_1$, except for a set of measure zero, then each row has at least one element of positive measure.

    But since we have more columns than rows, then, by the Pigeonhole Principle, there is at least some column with uncountably many disjoint sets of positive measure!
\end{sketch}

\begin{proof}
    We already saw that $\kappa$ should be regular, then, assume by a contradiction that $\kappa=\lambda^+$. By the previous theorem there is an Ulam $(\lambda^+,\lambda)$-matrix.

    Since the union of the sets in each row is all of $\lambda^+$, except for a set of measure $0$, then each row has at least one set of positive measure. However, by the Pigeonhole Principle, since there are more rows than columns, we can find a column with uncountably many disjoint sets of positive measure, contradicting $\sigma$-saturation, so $\kappa$ is a limit cardinal, implying it's weakly inaccessible.
\end{proof}

As a summary of all the nice work we have seen so far, we can state the following theorem:

\begin{theorem}{(Ulam)}{ulam}
    If there exists a measure on some set $S$, then either:
    \begin{itemize}
        \item there is a measure with an atom over $S$, and $\lvert S\rvert$ is $\geq$ the least inaccessible cardinal, or;
        \item there is an atomless measure on $\mf{c}$, making it $\geq$ the least weakly inaccessible cardinal.
    \end{itemize}
\end{theorem}

Therefore, the solution for the Set Theoretic Measure Problem is that: either we have a ``trivial'' measure on a monstrously large cardinal we can't even prove to exist in ZFC, or we have a ``non-trivial'' measure on $\mf{c}$, making it also comically big\footnote{how beautiful are God's creations}.

\begin{proof}
    If there is a measure on $S$, then either it's $2$-valued, and then if $\kappa\leq\lvert S\rvert$ is the least cardinal with a $2$-valued measure, by \cref{lem:measure_implies_kappa_add} it's measurable, so \cref{th:measurable_inaccessible} guarantees $\kappa$ is inaccessible; or it's atomless and, by \cref{th:atomless}, there is some atomless measure on a $\lambda\leq\mf{c}$. So, if $\kappa$ is the least cardinal with a measure in it, by \cref{lem:measure_implies_kappa_add} we have that $\kappa$ is real-valued measurable and, by \cref{th:real_valued_weakly_inacc}, $\kappa$ is weakly inaccessible, so $\mf{c}\geq\lambda\geq\kappa\geq\kappa_0$ is greater or equal than the least weakly inaccessible cardinal $\kappa_0$.
\end{proof}

\section{Compactness}

There are actually two natural ways to motivate the study of Compact Cardinals. The first, which we chose as our starting point, comes from the attempt to generalize the first order Compactness Theorem for infinitary logics. And the second comes from the attempt of generalizing Ramsey's Theorem. We will dedicate one section for each, since they're both interesting.

\subsection{Infinitary Logics}

As we said, another natural context where large cardinals appears are in the attempts of generalizing Compactness Theorem. Let's consider the Infinitary Logics $\mc{L}_{\kappa\lambda}$ with $\max\{\kappa,\lambda\}$ many variables and formulas allowing infinite conjunctions $\bigwedge_{\alpha<\eta}$ and disjunction $\bigvee_{\alpha<\eta}$, for any $\eta<\kappa$, and infinite quantifications $\forall_{\alpha<\eta}$ and $\exists_{\alpha<\eta}$, for any $\eta<\lambda$. The satisfaction relation is extended as usual.

In this notation, First Order Logic is just $\mc{L}_{\omega\omega}$, which, as we know, satisfies the Compactness Theorem. In order to generalize Compactness for $\mc{L}_{\kappa\lambda}$, let's first generalize finite satisfiability.

\begin{definition}{$\mu$-satisfiable}{mu_satisfiable}
    A set $\Phi$ of formulas in $\mc{L}_{\kappa\lambda}$ is \lasereye{$\mu$-satisfiable} if each $\Phi_0\subseteq\Phi$, with $\lvert\Phi_0\rvert<\mu$, is satisfiable.
\end{definition}

In this notation being finitely satisfiable is just being $\aleph_0$-satisfiable. Let's now generalize the Compactness Theorem for infinitary logics.

\begin{definition}{Weakly and Strongly Compact}{weakly_strongly_compact}
A cardinal $\kappa$ is \lasereye{weakly compact} (\lasereye{strongly compact}) iff any $\mc{L}_{\kappa\kappa}$-languages $L$ in a signature $S$ (with $\lvert S\rvert\leq\kappa$) satisfies that for every $\Phi\subseteq L^S$, we have that:
\vspace{-4pt}
$$\text{if $\Phi$ is $\kappa$-satisfiable, then $\Phi$ is satisfiable.}$$
\end{definition}

As a warm up, let's look at which properties our compact cardinals have. In fact, since the only tool we have are set of formulas in infinitary languages, many of the following proofs will be based on searching for neat set of formulas, which is a fun task that we strongly encourage you to try on your own first.

\begin{theorem}{}{weakly_compact_inaccessible}
    Every weakly compact cardinal $\kappa$ is inaccessible.
\end{theorem}

\begin{proof}
    Let's start with regularity. Assume by a contradiction that $\kappa=\bigcup X$, for some $X\subseteq\kappa$ with $\lvert X\rvert<\kappa$, and consider $S=\{c_\alpha:\alpha<\kappa\}\cup\{c\}$. Clearly $\lvert S\rvert\leq\kappa$, so we can apply compactness. Let
    $$\varphi=\bigvee_{\beta\in X}\bigvee_{\alpha<\beta}(c_\alpha=c)$$
    Since both $\lvert X\rvert,\beta$ are less than $\kappa$, the above disjunction is well-defined. Then, consider
    $$\Phi=\{c_\alpha\neq c:\alpha<\kappa\}\cup\{\varphi\}$$
    Now, if $\Phi_0\subseteq\Phi$ has $\lvert\Phi_0\rvert<\kappa$, we can find $\alpha<\kappa$ such that $(c_\alpha\neq c)\notin\Phi_0$, so we just take a model where both $c$ and $c_\alpha$ has the same interpretation, while satisfying $\Phi_0$. Then, by Compacteness we have that $\Phi$ is satisfiable, a contradiction. So $\kappa$ is regular.

    Now, for strong limit, assume by a contradiction that $2^\lambda\geq\kappa$ for some $\lambda<\kappa$. Let
    $$S=\{c_\alpha:\alpha<\lambda\}\cup\{0,1\},$$
    we want to express that $c_\alpha$ encodes a function in $2^\lambda$, so let
    $$\varphi:=\bigwedge_{\alpha<\lambda}(c_\alpha=0\vee c_\alpha=1).$$
    Now, the contradiction will come from the fact that we will say ``$c_\alpha\neq f$'' for every $f\in2^\lambda$, but still $c_\alpha$ itself will encode a function in $2^\lambda$. To do that consider, for each $f\in2^\lambda$:
    $$\psi_f:=\bigvee_{\alpha<\vee}(c_\alpha\neq f(\alpha))$$
    And let $\Phi=\{\varphi,0\neq1,\psi_f:f\in2^\lambda\}$. Now, if $\Phi_0\subseteq\Phi$ is such that $\lvert\Phi_0\rvert<\kappa\leq2^\lambda$, then there is some $f\in2^\lambda$ such that $\psi_f\notin\Phi_0$. Hence, if we let $c_\alpha$ be interpreted as $f$, we will have that $\Phi_0$ is satisfiable. Since $\lvert S\rvert\leq\kappa$, we can then apply compactness and obtain that $\Phi$ is satisfiable.

    However, if $M\vDash\Phi$, we can then define $g:\lambda\to2$ by
    $$g(\alpha)=\begin{cases}
        0,\text{ if }M\vDash(c_\alpha=0);\\
        1,\text{ if }M\vDash(c_\alpha=0).
    \end{cases}$$
    But since $\psi_g\in\Phi$, $c_\alpha\neq g(\alpha)$ for some $\alpha<\lambda$, a contradiction.
\end{proof}

\subsection{Łoś's Theorem for \texorpdfstring{$\mc{L}_{\kappa,\kappa}$}{Lkk}}


Now that we know that $\kappa$ being weakly compact ($C_w(\kappa)$), and hence strongly compact ($C_s(\kappa)$), is a Large Cardinal Property, a lot of natural questions arise. However, before diving into them, we need some model theoretic tools, mainly what is called ultraproducts.

In particular, we will ramble a little bit about Łoś's Theorem, which is a well known result on ultraproducts, but we will focus on it's generalization for infinitary logics. The proof is basically the same, but you use the fact we're quotienting by a $\kappa$-complete ultrafilter, which allows us to prove the $\bigwedge$ cases. So, since we're already proving this general version, I will devote this subsection to be a bit wordy on ultraproducts, then fell free to skip if you already know Łoś's Theorem and if you're convinced we can extend the same proof for infinitary logics.

\textcolor{red}{Rambling todo}

\subsection{Relation to Measurables}

With ultraproducts in hands, let's start showing that
$$C_s(\kappa)\implies M(\kappa),$$
which can be done using the following characterization of strongly compact cardinals.

\begin{theorem}{(Keisler-Tarski)}{kaisler_tarski}
    $\kappa$ is strongly compact iff every $\kappa$-complete filter can be extended to a $\kappa$-complete ultrafilter.
\end{theorem}

The reason why we need strongly compact here is because we will develop a language to talk about filters in $X$, so we will add one constant for each subset of $X$, meaning we will need at least $2^{|X|}$-many, which can eventually be greater than $\kappa$.

Intuitively, the sketch of the proof is to use the magical power of compactness to create a non-standard model with a ``ghostly" element $c$ of $X$ in our language, and then create our ultrafilter as the principal filter at $c$.

\begin{proof}
    $(\Rightarrow)$ Let $\mc{F}$ be a $\kappa$-complete filter on $X$, let $L$ be the $\mc{L}_{\kappa,\kappa}$-language
    $$L:=\{\in,c\}\cup\{c_A:A\subseteq X\}$$
    and let $M$ be the $L$-structure
    $$M:=(X\cup\mc{P}(X),\in,\{A:A\subseteq X\}).$$
    Now, to force the existence of a ghostly element $c$ that is inside every $A\subseteq X$ we let
    $$\Phi:=\mathrm{Th}(M)\cup\{c\in c_A:A\in\mc{F}\}$$
    then, given any $\Phi_0\subseteq\Phi$ with $\lvert\Phi_0\rvert=\lambda<\kappa$, at most $\lambda$-many $A_\alpha\in\mc{F}$ appears in $\Phi_0$. However, since $\mc{F}$ is $\kappa$-complete, we can let $c$ be interpreted as any element of $\bigcap_{\alpha<\lambda}A_\alpha\in\mc{F}$, so $\Phi_0$ is satisfiable and, therefore, by $\kappa$-compactness we have that $\Phi$ has a model $N$. We then let
    $$\mc{U}:=\{A\subseteq X:N\vDash(c\in c_A)\}$$
    and claim $U$ is such ultrafilter.

    In fact, if $A\in\mc{F}$, then $(c\in c_A)\in\Phi$ and, therefore, $N\vDash(c\in c_A)$, i.e., $A\in \mc{U}$ too, so $\mc{U}$ extends $\mc{F}$.

    $U$ is clearly a filter, to see it's an ultrafilter let $A\subseteq X$, then $M\vDash(c\in c_A)\vee(c\in c_{X\setminus A})$, and so $N$ too, meaning either $A\in \mc{U}$ or $X\setminus A\in U$, as desired.

    Lastly, we need to show $\mc{U}$ is $\kappa$-complete, and here is where the fact (not used so far) that we're working in an infinitary language enters. Let $\{A_\alpha\}_{\alpha<\lambda}\subseteq \mc{U}$, with $\lambda<\kappa$, and let
    $$A:=\bigcap_{\alpha<\lambda}A_\alpha.$$
    We want to show $c\in A$, and then we need to express infinite intersections in $\mc{L}_{\kappa,\kappa}$. This can be done as follows:
    $$M\vDash\forall x\left((x\in c_A)\leftrightarrow\bigwedge{\alpha<\lambda}(x\in c_{A_\alpha})\right)$$
    and since the RHS of $\leftrightarrow$ is satisfied for $x=c$ in $N$, then also is the LHS, i.e., $N\vDash(c\in c_A)$, so
    $$\bigcap_{\alpha<\lambda}A_\alpha\in \mc{U}$$
    as desired.

    $(\Leftarrow)$ \textcolor{red}{(Pendente -- Precisa de Los)}
\end{proof}

We can then prove what we wanted.

\begin{corollary}{(Erdős-Tarski)}{erdos_tarski}
    Any strongly compact cardinal is measurable.
\end{corollary}

\begin{proof}
    The idea is to consider the following filter on $\kappa$:
    $$\mc{F}=\{A\subseteq\kappa:\lvert\kappa\setminus A\rvert<\kappa\}$$
    because, given any extension $\mc{U}$ of $\mc{F}$, since $\kappa\setminus\{a\}\in\mc{F}\subseteq\mc{U}$ for any $a\in\kappa$, we can't have $\{a\}\in\mc{U}$. We then only need to guarantee that $\mc{F}$ is $\kappa$-complete and then, by Keisler-Tarski's Theorem we can extend it to a $\kappa$-complete ultrafilter on $\kappa$ and, by \cref{def:measurable_cardinal}, $\kappa$ is measurable.

    In fact, since $\kappa$ is regular, given $\{A_\alpha\}_{\alpha<\lambda}\subseteq\mc{F}$, with $\lambda<\kappa$, we have $\lvert\kappa\setminus A_\alpha\rvert<\kappa$, so, by regularity
    $$\left\lvert\bigcup_{\alpha<\lambda}(\kappa\setminus A_\alpha)\right\rvert=\left\lvert\kappa\setminus\rp{\bigcap_{\alpha<\lambda} A_\alpha}\right\rvert<\kappa$$
    meaning $\bigcap_{\alpha<\lambda}A_\alpha\in\mc{F}$, i.e., $\mc{F}$ is $\kappa$-complete.
\end{proof}

The next natural question is where weakly compact cardinals sit in relation to measurables. This is what the next Theorem, also due to Erdős and Tarski, answers.

\begin{theorem}{(Erdős-Tarski)}{erdos_tarski}
    Any measurable cardinal is weakly compact.
\end{theorem}

\begin{proof}
    \textcolor{red}{Pendent}
\end{proof}

\subsection{Elementary Submodels Technique}

Maybe surprisingly, there's another, seemly unrelated, approach to compact cardinals which is also very natural. As we did for the previous cardinals\footnote{even though we're talking about the same cardinal as before}, let's keep this motivational presentations of such cardinals and ramble a lit bit about Infinite Ramsey's Theory.

I will assume you at least know the finite version of Ramsey's Theorem in order to motivate the infinite version but, even though you've already heard a lot about it, trust me, there will be plenty of other cools stuffs you probably don't know. Mainly, I will now rant considerably on one of the most interesting proof techniques model theory gave us, called Elementary Submodels Technique, which I will introduce by solving one really hard topology problem, and by giving a combinatorics-free proof of the $\Delta$-system lemma! We will later refine it to prove Erdős-Rado theorem.



\subsection{Infinite Ramsey's Theory}